News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-23 rupnik2015.tex
system overview

The second step is to match our computed clusters to events. Since our similarity measure and clustering
may not be perfect, we need to find the ``best" correspondence between clusters and events. In this paper, we test several different algorithms for both steps for performance in an integrated system. In particular, we test our approach within an online system for detection of world events, called Event Registry\cite{Leban2014W, Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.

Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguation and entity linking of the concepts is done using Wikipedia as the main knowledge base.
   The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event.

@@

The paper is organized as follows: we first provide an overview of the system as a whole (Section ), followed by details on each of the two steps described above. We first introduce several approaches to measures of cross-lingual similarity (Section ) and then the algorithms for matching clusters to events.  To improve readability, we describe relevant related work for each module in its corresponding Section.
Finally, we present and interpret the experimental results and finally give conclusions and describe promising future directions.
%
%\section{Introduction}
%
%
%{\oldText{Taken from XLite paper}
%Document retrieval is a well-established problem in data mining. There have been a large number of different approaches put forward in the literature. In this paper we concentrate on a specific setting: multi-lingual corpora. As the availability of multi-lingual documents has exploded in the last few years, there is a pressing need for automatic cross-lingual processing tools.
%
%The prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2012, the percentage of English articles has dropped to 14\%. In this context, we look at how to find similar documents across languages. In particular, we do not assume the availability of machine translators, but rather try to frame the problem such that we can use well-established machine learning tools designed for monolingual text-mining tasks. This work represents the continuation of previous work \cite{nips, iti, nips2, sikdd} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. We look at a specific aspect of this problem: the distribution of articles across languages in Wikipedia is not uniform. While the percentage of English articles has fallen, English is still the largest language and one of \emph{hub} languages which not only have an order of magnitude more articles than other languages, but also many comparable articles with most of the other languages.
%
%When doing document retrieval, we encounter two quite different situations. If for example, we are looking for a German article comparable to an English article, there is a large alignment between the document corpora (given by the intersection in articles in the two languages) making the problem well-posed. If, however, we look for a relevant Slovenian article given a Hindi article, the intersection is small, making the problem much harder. Since almost all languages have a large intersection in articles with \emph{hub} languages, the question we ask in this paper is: can we exploit hub languages to perform better document retrieval between non-hub languages? A positive answer would vastly improve cross-lingual analysis between less represented languages. In the following section, we introduce our representation followed by our experiments, which also shed light on the structural properties of the multilingual Wikipedia corpus.
%}

\section{System Overview}

Event registry $\Rightarrow$ Monolingual events $\Rightarrow$ features (semantic + CCA) $\Rightarrow$ linked events (crosslingual)

@@

\section{Cross-lingual Event Linking}

The main application on which we test the above similiarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

\begin{comment}
One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''.

Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguated and entity linking of the concepts is done using Wikipedia as the main knowledge base. The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster.

In short, the article clustering works as follows (details are available in \cite{Brank2014}). First we create a new document by joining the title and the body of the article. We then tokenize the document, remove the stop words and stem the words. The remaining words are represented as a vector in the vector-space model. Each value in the vector is additionally normalized using TF-IDF. After normalization, we use the vector to find the most similar existing cluster. To be able to compare the vector with a cluster we first need a vector representation for each of the clusters. This is done by computing a centroid vector of the cluster by simply averaging the vectors of the articles in the cluster. By computing the cosine similarity of the article vector with each of the centroid vectors we then find the cluster with the highest similarity. If the highest similarity is above the selected threshold (we used 0.4 in our experiments), the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

In order to treat the cluster as describing an event in Event Registry, the cluster has to grow above a certain number of articles (the number depends on the language). This condition ensures that only content that receives a sufficient coverage in the media is treated as an event. Once a new event is identified, a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.
\end{comment}

@@

The problem in the second phase of the pipeline is to identify the clusters (given?) describing the same events. Each article $a_i$ is written in a language $l$, where $l \in L = \{l_1,l_2,...,l_k\}$. For each language $l$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{l_a}$, we would like to identify clusters $c_j \in C_{l_b}$ (where $l_a \neq l_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{l_a}$ and $C_{l_b}$ respectively.

@@

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are very few service that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that provides cross-lingual cluster linking is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\cite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists.

A system that is significantly different, but worth mentioning is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.

>>>>

The second step is to match our computed clusters to events. Since our similarity measure and clustering
may not be perfect, we need to find the ``best" correspondence between clusters and events. In this paper, we test several different algorithms for both steps for perfomance in an integrated system. In particular, we test our approach within an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.

Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguated and entity linking of the concepts is done using Wikipedia as the main knowledge base.
   The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event.

@@

The paper is organized as follows: we first provide an overview of the system as a whole (Section ), followd by details on each of the two steps described above. We first introduce several approaches to measures of cross-lingual similarity (Section ) and then the algorithms for matching clusters to events.  To improve readability, we describe relevant related work for each module in its corresponding Section.
Finally, we present and interpret the experimental results and finally give conclusions and describe promising future directions. 
%
%\section{Introduction}
%
%
%{\oldText{Taken from XLite paper}
%Document retrieval is a well-established problem in data mining. There have been a large number of different approaches put forward in the literature. In this paper we concentrate on a specific setting: multi-lingual corpora. As the availability of multi-lingual documents has exploded in the last few years, there is a pressing need for automatic cross-lingual processing tools.
%
%The prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2012, the percentage of English articles has dropped to 14\%. In this context, we look at how to find similar documents across languages. In particular, we do not assume the availability of machine translators, but rather try to frame the problem such that we can use well-established machine learning tools designed for monolingual text-mining tasks. This work represents the continuation of previous work \cite{nips, iti, nips2, sikdd} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. We look at a specific aspect of this problem: the distribution of articles across languages in Wikipedia is not uniform. While the percentage of English articles has fallen, English is still the largest language and one of \emph{hub} languages which not only have an order of magnitude more articles than other languages, but also many comparable articles with most of the other languages.
%
%When doing document retrieval, we encounter two quite different situations. If for example, we are looking for a German article comparable to an English article, there is a large alignment between the document corpora (given by the intersection in articles in the two languages) making the problem well-posed. If, however, we look for a relevant Slovenian article given a Hindi article, the intersection is small, making the problem much harder. Since almost all languages have a large intersection in articles with \emph{hub} languages, the question we ask in this paper is: can we exploit hub languages to perform better document retrieval between non-hub languages? A positive answer would vastly improve cross-lingual analysis between less represented languages. In the following section, we introduce our representation followed by our experiments, which also shed light on the structural properties of the multilingual Wikipedia corpus.
%}

\section{Pipeline}

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.  Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguated and entity linking of the concepts is done using Wikipedia as the main knowledge base. An online clustering algorithm is then applied on the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster.

The processing is done independently for each language and the output is a clustering of articles in each language, with each cluster ideally representing an event, which are defined by topically, geographically and temporally. The goal of this paper is present approaches to  connecting the clusters/events across languages.

Our approach is based on first defining a similarity function between articles in different languages (Section~\ref{}) and then we present several efficient approaches to linking the clusters based on features (Section~\ref{}). First, however we describe the pre-processing on the input articles which we then take as input.

The system is designed to handle streams and so we describe how a document which enters the stream is processed.   A new document is first tokenized, stop words are removed and words are stemmed. 
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{} for the definition). Since each article is tagged with its language, cosine similiarity  is used to find the modst similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

The clusters, we consider must have a sufficient number of articles (which is a language dependent paramater). This is to ensure that we consider events. Once an event is identified(i.e. a cluster of sufficient size), a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

 Further details about the system can be found in \cite{}, while our full pipeline is shown in Figure~\ref{}.

\begin{comment}
In short, the article clustering works as follows (details are available in \cite{}). First we create a new document by joining the title and the body of the article. We then tokenize the document, remove the stop words and stem the words. The remaining words are represented as a vector in the vector-space model. Each value in the vector is additionally normalized using TF-IDF. After normalization, we use the vector to find the most similar existing cluster. To be able to compare the vector with a cluster we first need a vector representation for each of the clusters. This is done by computing a centroid vector of the cluster by simply averaging the vectors of the articles in the cluster. By computing the cosine similarity of the article vector with each of the centroid vectors we then find the cluster with the highest similarity. If the highest similarity is above the selected threshold (we used 0.4 in our experiments),

In order to treat the cluster as describing an event in Event Registry,

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.
\end{comment}
%Event registry $\Rightarrow$ Monolingual events $\Rightarrow$ features (semantic + CCA) $\Rightarrow$ linked events (crosslingual)

@@

\section{Cross-lingual Event Linking}
The main application we test the above simliarity on  is cross-lingual event linking. In online media streams - particularly new articles - there is often duplication of reporting, different viewpoints or opinions, all centering around a single event.  Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015. (Sports example)

Informally, events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry\cite{}. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to find suitable matchings between clusters with articles in different languages.

@@

The problem in the second phase of the pipeline is to identify the clusters describing the same events. Each article $a_i$ is written in a language $l$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_k\}$. For each language $\ell$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{\ell_a}$, we would like to identify clusters $c_j \in C_{\ell_b}$ (where $\ell_a \neq \ell_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{\ell_a}$ and $C_{\ell_b}$ respectively. Furthermore, we iterate over all the languages we consider. That is, given a cluster in language $\ell_i$, we look for matching clusters in language $\ell_j$ for $i\neq j$  in a pairwise fashion.

Furthermore, the matching of clusters is a \emph{generalized matching}. We cannot assume that there is only one cluster per language per event, nor can we assume complete coverage - that there exists at least one cluster per event in every language. This implies that we cannot make any assumptions on the matching, e.g. one-to-one, complete, etc. This excludes the use of standard weighted bipartite matching type of algortihms for this problem.

@@

TODO
